---
title: "Stout Case Study 2: Data Manipulation and Visualization"
author: "Theresa Pham"
date: "12/24/2022"
output: html_document
---

```{r setup, echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


***Read and look at data***
```{r}
orders <- read.csv("casestudy.csv")
glimpse(orders)
head(orders)
```


## Data Manipulation
We will extract and calculate information for each year, so we will find how many years there are in the data set. 
```{r}
table(orders$year)
```


For each prompt, we should obtain results for each year applicable. 

**Total revenue for the current year**
```{r}
Total_Rev <- orders %>%
  group_by(year) %>%
  summarise(total_rev = sum(net_revenue))
Total_Rev
```


**New Customer Revenue** (new customers not present in previous year only)
```{r}
custs2015 <- orders %>%
  filter(year == 2015) 
custs2016 <- orders %>%
  filter(year == 2016)
custs2017 <- orders %>%
  filter(year == 2017)

# all customers in 2015 are "new" for 2015
new2015 <- custs2015

# new customers in 2016
new2016 <- anti_join(custs2016, custs2015, by="customer_email")

# new customers in 2017
new2017 <- anti_join(custs2017, custs2016, by="customer_email")

# create new new data frame with new customers only
new_orders <- bind_rows(new2015, new2016, new2017)

New_Cust_Rev <- new_orders %>%
  group_by(year) %>%
  summarise(new_cust_rev = sum(net_revenue))
New_Cust_Rev
```

Note that we do not have data on years before 2015, thus we cannot accurately determine whether customers are new in 2015.

**Existing Customer Growth** (Revenue of existing customers for current year minus Revenue of existing customers from the previous year)
```{r}
# existing customers = all current customers - new customers

existing_rev <- Total_Rev %>%
  left_join(New_Cust_Rev, by="year") %>%
  mutate(existing_rev = total_rev - new_cust_rev) %>%
  mutate(prev_existing_rev = lag(existing_rev))
existing_rev

Existing_Cust_Growth <- existing_rev %>%
  mutate(existing_cust_growth = existing_rev - prev_existing_rev) %>%
  select(year, existing_cust_growth)
Existing_Cust_Growth
```
Note that revenues may be off again due to missing data from year before 2015. We assumed all 2015 customers to be "new", thus there is no existing customer revenue for that year.


**Revenue lost from attrition** (revenue lost from lost customers)
```{r}
# We are assuming no customers lost in 2015

# lost customers in 2016 
lost2016 <- anti_join(custs2015, custs2016, by="customer_email")

# lost customers in 2017
lost2017 <- anti_join(custs2016, custs2017, by="customer_email")

# create new new data frame with lost customers only
lost_custs <- bind_rows(lost2016, lost2017)

Attrition_Rev_Lost <- lost_custs %>%
  group_by(year) %>%
  summarise(attrition_rev_lost = sum(net_revenue)) %>%
  rename(prev_year = year) %>%
  mutate(year = prev_year + 1) %>%
  add_row(prev_year = 2014, year = 2015, attrition_rev_lost = 0) %>%
  arrange(prev_year)

col_order <- c("prev_year", "year", "attrition_rev_lost")
Attrition_Rev_Lost <- Attrition_Rev_Lost[, col_order]
Attrition_Rev_Lost
```
Since we do not have data from 2014, we assume no customers were lost in 2015. 

**Existing Customer Revenue Current Year**
```{r}
Current_Existing_Rev <- existing_rev %>%
  select(year, existing_rev)
Current_Existing_Rev
```

**Existing Customer Revenue Prior Year**
```{r}
Prev_Existing_Rev <- existing_rev %>%
  select(year, prev_existing_rev)
Prev_Existing_Rev
```

**Total Customers Current Year**
```{r}
Total_Custs <- orders %>%
  group_by(year) %>%
  summarise(total_custs = n())
Total_Custs
```

**Total Customers Previous Year**
```{r}
Prev_Total_Custs <- Total_Custs %>%
  mutate(prev_total_custs = lag(total_custs)) %>%
  select(year, prev_total_custs)
Prev_Total_Custs
```

**New Customers** 
```{r}
New_Custs <- new_orders %>%
  group_by(year) %>%
  summarise(new_custs = n())
New_Custs
```

**Lost Customers**
```{r}
Lost_Custs <- lost_custs %>%
  mutate(current_year = year + 1) %>%
  group_by(current_year) %>%
  summarise(lost_custs = n()) %>%
  rename(year = current_year) %>%
  add_row(year = 2015, lost_custs = 0) %>%
  arrange(year)
  
Lost_Custs
```


## Visualizations and Observations

**Grouped Bar Chart Comparing Lost, New, and Existing Customer Count**
```{r}
# Data frame with existing, new, and lost customer count
Existing_Custs <- Total_Custs %>%
  full_join(New_Custs) %>%
  mutate(existing_custs = total_custs - new_custs)
#Existing_Custs

custs_df <- Existing_Custs %>% full_join(Lost_Custs, by='year')
custs_df
```
```{r}
custs_df_long <- custs_df %>%
  pivot_longer(cols = 3:5, 
               names_to = "cust_type", 
               values_to = "cust_count") 

custs_df_long$cust_type <- factor(custs_df_long$cust_type, levels = c('lost_custs', 'new_custs', 'existing_custs'))

custs_df_long$year <- factor(custs_df_long$year, levels = c('2017', '2016', '2015'))

custs_df_long %>%
  ggplot(aes(x=year, y=cust_count, fill=cust_type)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip() +
  ggtitle("Lost, New, and Existing Customers per Year") +
  labs(x="Year", y="Customer Count", fill="Customer Type") + 
  scale_fill_discrete(labels=c("Lost Customers", "New Customers", "Existing Customers")) 
```

Observations: Although we only have data on three years and assume that all customers in 2015 are "new", we can observe that the number of existing, or returning, customers is far below the number of lost customers, thus customer retention is not well. We can also see that the number of lost and and new customers are both increasing in the latter two years. 


**Line Graph of Total Revenue Over Time**

```{r}
Total_Rev %>%
  ggplot((aes(x=year, y=total_rev))) +
  geom_line(color="red") +
  ggtitle("Total Revenue Over Time") +
  labs(x="Year", y="Total Revenue") + 
  scale_x_continuous(breaks = seq(2015, 2017, 1)) + 
  ylim(0, 32500000)
```

Observations: The total revenue does not fluctuate drastically within the three recorded years, but we can notice a dip in revenue in 2016. 


**Stacked Bar Chart of New and Existing Customer Revenue**
```{r}
# Data frame with total, existing, and new revenue
rev_list <- list(Total_Rev, Current_Existing_Rev, New_Cust_Rev)

rev_df <- rev_list %>% reduce(full_join, by = 'year')
rev_df
```

```{r}
rev_df_long <- rev_df %>%
  pivot_longer(cols = 3:4, 
               names_to = "rev_type", 
               values_to = "rev_amount") 
rev_df_long

rev_df_long$rev_type <- factor(rev_df_long$rev_type, levels = c('new_cust_rev', 'existing_rev'))

rev_df_long %>%
  ggplot(aes(x=year, y=rev_amount, fill=rev_type)) + 
  geom_bar(stat="identity", position="stack") +
  ggtitle("Revenue from Existing and New Customers per Year") +
  labs(x="Year", y="Revenue", fill="Revenue Type") + 
  scale_fill_brewer(labels=c("New Customers", "Existing Customers"), palette = "Set2") + 
  geom_text(aes(x=year, y=rev_amount, label=rev_amount, group=rev_type), 
            position=position_stack(vjust = .5), colour="white")
```

Observations: The majority of total revenue is made up of new customers every year, and 2017 has the highest total revenue out of all three years.

